We're going to work with some IMDB data. First we'll load it and subset it to four genres just to keep things simple. We're also going to use `ddply` to calculate the median rating for each year. We're going to ask: Does the distribution of ratings change through time across these four genres?

```{r}
library(plyr)
load("imdb.rda")
genres <- c("Action", "Drama", "Romance", "Comedy")
imdb <- imdb[imdb$genre %in% genres, ]
d <- ddply(imdb, c("genre", "year"), summarize, median_rating = median(imdb_rating))
```

Now, let's plot out the median ratings by year with panels for each genre. We'll use `par(mfrow)` this time:

```{r}
par(mfrow = c(2, 2))
for(i in 1:4) {
  dat <- subset(d, genre == genres[i])
  with(dat, plot(year, median_rating))
  mtext(genres[i])
}
```

In this case we can remove the inner margin space and remove redundant axes:

```{r}
par(mfrow = c(2, 2))
par(mar = c(0, 0, 0, 0),
  oma = c(5, 5, .5, .5))
for(i in 1:4) {
  dat <- subset(d, genre == genres[i])
  with(dat, plot(year, median_rating, axes = FALSE, ylim = range(d$median_rating)))
  mtext(genres[i], line = -2)
  box(col = "grey50")
  if(i %in% c(1, 3)) axis(2, col = "grey50")
  if(i %in% c(3, 4)) axis(1, col = "grey50")
}
  mtext("Year", side = 1, outer = TRUE, line = 2)
  mtext("Rating", side = 2, outer = TRUE, line = 2)
```

Now, we'll use `layout` to make a plot with a long plot on the top and two narrower plots on the bottom. We'll plot a time series of year against IMDB rating on the top, a histogram of ratings on the bottom left, and a histogram of votes on the bottom right.

```{r}
layout(rbind(c(1, 1), c(2, 3)))
layout.show(3)
par(mar = c(3, 2, 0, 0), oma = c(1, 1, 1, 1))
par(cex = 0.7)
with(imdb, plot(year, budget))
hist(imdb$imdb_rating)
hist(imdb$imdb_votes)
```

Now, let's try using `split.screen`. We'll come up with an arbitrary layout for 3 panels, create a `.csv` file containing the positions, and plot 3 panels.

```{r}
m <- read.csv("split.screen.csv")
split.screen(as.matrix(m))
screen(1)
par(mar = c(0,0,0,0))
plot(1)
screen(2)
par(mar = c(0,0,0,0))
plot(1)
screen(3)
par(mar = c(0,0,0,0))
plot(1)
close.screen(all = TRUE)
```
