NOTE THAT THIS ASSIGNMENT IS NOT COMPLETE YET. 
PLEASE WAIT TO START WORKING ON IT!

# Assignment 3: `plyr` and `reshape`

For this assignment include all your R code in this R Markdown file to support your answer. Your code goes in the code "chunks" below. If you open this file in RStudio then you can run all the R code by clicking on the `Knit HTML` button. You can experiment with code in the R console and then paste your code into this document, or write it directly in this document and send it over to the console with the keyboard shortcut `command+enter` or `control+enter`.

Your professor (the expert in Fire Pokemon) has been denied tenure (what'd you expect, he studies Pokemon...) and has quit in disgrace.
Fortunately for you another Biology professor (one that studies, you know, actual living things) has agreed to take you on as a student. 
She has just discovered [BIRDD](http://bioquest.org/birdd/morph.php) makes morphological data from Galapagos finches freely available online. 
This is great because you don't have money for a field season. 
There must be a dissertation chapter in this data somewhere! 
Time to tidy up the data and see what gems you can find.

1. Load the morphological data provided on the BIRDD webpage into the R environment.
Do not display this code in the knitted document.

```{r}

```

2. Take a look at the data
  - Print the first 6 rows of the data frame.
  - Use a command to determine the dimensions of the data frame (number of rows, number of columns?).
  - Is the data in long or wide format?

```{r}

```

3. When manipulating a larger data frame like this it is often useful to develop your code using just a subset.
One way to do this is by indexing using the square brackets `[ ]`. 
Using the example below as a guide, create a subset of your data frame that has only the first 100 rows and all columns.  

```{r}
# the general form is: 
# my.df_short <- my.df[r,c] 
# where r is the row or rows and c is the same for columns. Leaving a slot blank will take all rows or columns
```

4. [Some plyr exercises. Re-calculate the data for some of the tables on the BIRDD website?]

```{r}

```

5. [Reshape the data into long format using `melt` and save for use in the ggplot assignment.]

```{r}

```
