NOTE THAT THIS ASSIGNMENT IS NOT COMPLETE YET. 
PLEASE WAIT TO START WORKING ON IT!

# Assignment 3: `plyr` and `reshape2`

Your professor (the expert in Fire Pokemon) has been denied tenure (what did
you expect, he studies Pokemon...) and has quit in disgrace. Fortunately for
you, another biology professor (one that studies, you know, actual living
things) has agreed to take you on as a student. She has just discovered that
[BIRDD](http://bioquest.org/birdd/morph.php) makes morphological data from
Galapagos finches freely available online. This is great because you don't
have money for a field season. There must be a dissertation chapter in this
dataset somewhere! Time to tidy up the data and see what gems you can find.

First, load the `plyr` and `reshape2` libraries:

```{r}
library(plyr)
library(reshape2)
```

1.  Load the morphological data provided on the 
    [BIRDD webpage](http://bioquest.org/birdd/morph.php) into a `data.frame`
    named `morph`. (Hint: it's probably easier to download the Excel file and
    convert it to a `.csv`.) Once you've loaded the `.csv` file into `R`, make
    all the column names lowercase with the `tolower()` function. Do not display
    this code chunk in the knitted document. 


```{r, echo=FALSE}
# Bonus `R` tip: you can download a file through `R` like this:
# download.file("http://bioquest.org/birdd/Morph_for_Sato.xls", destfile = "Morph_for_Sato.xls")
# now go look in your R working directory for the file
```

```{r}

```

2.  Take a look at the data. First, print the first 6 rows of the data frame.
    (Use the `head()` command.) Then, use a command to determine the dimensions
    of the data frame (number of rows, number of columns). Finally, is the data
    (the morphological measurements themselves) in long or wide format?

```{r}

```

    Just to make things simpler, we're going to remove the columns `n.ubkl`,
    `bodyl`, `plumage`. To do that, just run these lines:

```{r}
morph$n.ubkl <- NULL
morph$bodyl <- NULL
morph$plumage <- NULL
```

3.  If you scroll down on the BIRDD webpage where you got the data file you'll
    see a table of summary statistics. Using `ddply()`, we're going to recreate
    parts of the first table (Statistics by Island, Species for the
    Genovesa/Tower island). 

    First, subset the data to only include the `islandid` `Gnov_Twr` and call
    this new `data.frame` `morph_gt`.

    Then, use `ddply()` along with `summarize` to calculate the mean and
    standard deviation for beak height (`beakh`) for each species (`speciesid`)
    for this new object `morph_gt`. Within your `ddply` line, tell `mean` and
    `sd` to remove the `NA` values.

    Check a couple of the species with the website. Do they match?

```{r}

```

4.  There are a lot of `NA` values in this dataset. Let's figure out how many
    `NA`s there are for each of these measurements. One quick way to do this is
    by `melt`ing the measurements into long format and using `ddply`.

    First, `melt` `morph_gt` so that you have an ID row for `speciesid`,
    `subspid`, and `sex`. Keep all the remaining measured columns to the right.
    Name your new long-format `data.frame` object `morph_gt_melt`.

    Hint: since we don't want the first 6 columns in our melted dataset, we can
    remove them with `morph_gt[ , -c(1:6)]`. Do that first.

```{r}

```

5.   Now, use `ddply` to count the number of `NA` values for each combination of
    `speciesid`, `subspid`, and `sex`. Hint: you'll need to combine `is.na()`
    and `sum()`. Test out what happens if you run `sum(c(TRUE, TRUE))` vs.
    `sum(c(TRUE, FALSE))` and note that `is.na()` returns `TRUE` or `FALSE`.

```{r}

```

6.  Now, take `morph_gt_melt` and make it wide again using the `dcast`
    function. Maintain ID columns for `speciesid`, `subspid`, and `sex`. Since
    we'll be condensing multiple rows of data, use the argument `fun.aggregate`
    to calculate the mean value (make sure to exclude `NA`s). (Hint: see the
    example in class for how to pass `na.rm = TRUE` to `dcast`.)

```{r}

```

Notice that you've just used `dcast` to create the summaries on the
[BIRDD webpage](http://bioquest.org/birdd/morph.php). Combining `melt` and 
`dcast` is another way to do what you did with `ddply` in step 3.

Make sure your whole script runs when you click the "Knit HTML" button in
Rstudio.
