% An introduction to the `plyr` package
% 2013-10-07
% Sean Anderson, sean@seananderson.ca

For these exercises we'll use the `french_fries` dataset from the `reshape2`
package.

```{r}
library(reshape2)
head(french_fries)
```

Did the mean rancid flavour change through time?

```{r}
rancid <- ddply(french_fries, "time", summarize, mean_rancid = mean(rancid,
    na.rm = TRUE))
plot.default(rancid)
```

What is the mean rancid rating for each time-treatment combination?

```{r}
ddply(french_fries, c("time", "treatment"), summarize, mean_rancid =
  mean(rancid, na.rm = TRUE))
```

How many subjects are there for each treatment? Use `ddply` with the
`summarize` function to find out.

```{r}
ddply(french_fries, "treatment", summarize, n_subject =
  length(unique(subject)))
```

Say you wanted to standardize the buttery index within each subject. You could
do this by dividing the buttery rating within each subject by the maximum
buttery rating that subject gave throughout the study. You would come up with
a "percentage buttery" value. You can do this with the `transform` function
within `ddply`. Try that now: 

```{r}
ff2 <- ddply(french_fries, "subject", transform, 
  buttery_percentage = 100 * buttery / max(buttery))
head(ff2)
```

