% An introduction to the `plyr` package
% 2013-10-09
% Sean Anderson, sean@seananderson.ca

For these exercises we'll use the `french_fries` dataset from the `reshape2`
package.

```{r}
library(reshape2)
library(plyr)
head(french_fries)
```

Did the mean rancid flavour change through time?

```{r}
ff <- ddply(french_fries, "time", summarize, mean_rancid_rating = mean(rancid, na.rm = TRUE))
plot(ff)
plot.default(ff)
```

What is the mean rancid rating for each time-treatment combination?

```{r}
ff2 <- ddply(french_fries, c("time", "treatment"), summarize, mean_rancid = mean(rancid, na.rm = TRUE), sd_rancid = sd(rancid, na.rm = TRUE))
```

How many subjects are there for each treatment? Use `ddply` with the
`summarize` function to find out.

```{r}
ff5 <- ddply(french_fries, "treatment", summarize, length(unique(subject)), mean(buttery))
max(as.numeric(as.character(french_fries$subject)))
```

Say you wanted to standardize the buttery rating within each subject. You
could do this by dividing the buttery rating within each subject by the
maximum buttery rating that subject gave throughout the study. You would come
up with a "percentage buttery" value. You can do this with the `transform`
function within `ddply`. Try that now: 

```{r}
ff3 <- ddply(french_fries, "subject", transform, percentage_buttery = buttery / max(buttery))
```

Some examples with the `baseball` dataset where we return character objects:

```{r}
# Get first listed player id for each team:
b1 <- ddply(baseball, "team", summarize, first_id = id[1])
# split by team, year
# return id with most hr (there might be problems here if multiple players have
# the same max home runs)
b2 <- ddply(baseball, c("team", "year"), summarize, max_hr = id[hr == max(hr)])
```
