% An introduction to the `reshape2` package
% 2013-10-09
% Sean Anderson, sean@seananderson.ca

First, load the `reshape2` package and the `airquality` dataset: 

```{r}
library(reshape2)
names(airquality) <- tolower(names(airquality))
head(airquality)
```

What happens if you just run the `melt` function on the `airquality` data?

```{r}
a <- melt(airquality)
head(a)
```

Let's melt the `airquality` dataset but maintain a row of data for `month` and
`day`:

```{r}
a <- melt(airquality, id.vars = c("month", "day"))
```

We can also control the variable and value column names with `variable.name`
and `value.name` arguments. Try changing these column names while you run the
`melt` function:

```{r}
a2 <- melt(airquality, id.vars = c("month", "day"), variable.name =
  "climate_var", value.name = "climate_val")
```

Now let's try making the long data wide again with the `dcast` function. First,
let's go back to the default column names:

```{r}
a2 <- melt(airquality, id.vars = c("month", "day"))
```

Now, let's try and re-create the original wide data frame with `dcast` and
compare it to the original data frame:

```{r}
a2_wide <- dcast(a2, month + day ~ variable, value.var = "value")
```

What happens if you use `dcast` but don't specify day as an ID variable?

```{r}
a3_wide <- dcast(a2, month ~ variable, value.var = "value")
```

This time aggregate the multiple days by taking a `mean` value:

```{r}
a4_wide <- dcast(a2, month ~ variable, value.var = "value", fun.aggregate =
  mean, na.rm = TRUE)
```

OK, now try this on a different dataset. The `reshape2` package contains a
dataset called `french_fries` that describes the results from a study that
tested the effect of frying oils on French fry taste.

Try `melt`ing the data while keeping ID columns for `time`, `treatment`,
`subject`, and `rep`.

```{r}
ff_long <- melt(french_fries, id.vars = c("time", "treatment", "subject",
    "rep"))
```

And then `cast` the data frame you created back into wide format: 

```{r}
ff_wide <- dcast(ff_long, time + treatment + subject + rep ~ variable,
  value.var = "value")
```
