# Assignment 1: Introduction to R

For this assignment include all your R code in this R Markdown file to support your answer. Your code goes in the code "chunks" below. If you open this file in RStudio then you can run all the R code by clicking on the `Knit HTML` button. You can experiment with code in the R console and then paste your code into this document, or write it directly in this document and send it over to the console with the keyboard shortcut `command+enter` or `control+enter`.

Your professor (an expert in Fire Pokemon) has just returned from Vermillion City. He has been busy collecting data on traits of herds of [Charmanders](http://pokemon.wikia.com/wiki/Charmander). The professor, hearing you are in the new Data Wrangling course asks for you help summarize his findings.

1. Load the `charmander.csv` data file into the R environment. Don't display this code in the knitted document.

```{r}

```

2. Plot male and female Charmander mass.

```{r}

```

3. Ecologists are often accused of Physics envy, so pretend the Charmanders are [Voltorbs](http://pokemon.wikia.com/wiki/Voltorb)(spheres) and make a scatter plot of their mass vs height. Add a blue line through these data.

[DO THEY KNOW HOW TO FIT AND EXTRACT A REGRESSION LINE YET?]

```{r}

```

4. Re-plot the data and this time: (1) do not plot the axes with the `plot()` command, (2) add the axes with the `axis()` command with tick marks and labels at [FILL IN SOMETHING HERE], and (3) change the axis labels to something other than the default.

```{r}

```

5. Add a panel label `(a)` in the top left corner. Add a legend in the top right... [THERE ISN'T MUCH TO ADD A LEGEND ABOUT YET]

```{r}

```
