# Assignment 2: Basic Graphics

For this assignment include all your R code in this R Markdown file to support your answer. Your code goes in the code "chunks" below. If you open this file in RStudio then you can run all the R code by clicking on the `Knit HTML` button. You can experiment with code in the R console and then paste your code into this document, or write it directly in this document and send it over to the console with the keyboard shortcut `command+enter` or `control+enter`.

Your professor (an expert in Fire Pokemon) has just returned from Vermillion City. He has been busy collecting data on traits of herds of [Charmanders](http://pokemon.wikia.com/wiki/Charmander). The professor, hearing you are in the new Data Wrangling course asks for you help summarize his findings.

1. Load the `charmander.csv` data file into the R environment. Don't display this code in the knitted document.

```{r}

```

2. Plot male and female Charmander mass. Choose two colour-blind safe colours.

```{r}

```

3. Ecologists are often accused of Physics envy, so pretend the Charmanders are [Voltorbs](http://pokemon.wikia.com/wiki/Voltorb)(spheres) and make a scatter plot of their mass vs height. 

```{r}

```

4. Re-plot the data and this time: (1) do not plot the axes with the `plot()` command, (2) add the axes with the `axis()` command with tick marks and labels at [FILL IN SOMETHING HERE], and (3) change the axis labels to something other than the default.

```{r}

```

5. Add a legend in the top right two the charmander data. Have males and females have different style points. Add text directly to the graph identifying the smallest and the largest charmanders. Change the ticks on the graph to face inward and space them every 5 units.

```{r}

```
