An introduction to the `ggplot2` package
===================


```{r load-ggplot}
library(ggplot2)
# is your ggplot2 package 0.9.0 or greater?
sessionInfo()
```

Let's make a data frame and plot it with ggplot to get a feel for the syntax
```{r intro-code-from-slides}
# Introductory code from slides:
d <- data.frame(x = c(1:10, 1:10), 
                y = runif(20), 
                group1 = rep(gl(2, 5, labels = c("a", "b")), 2), 
                group2 = gl(2, 10))
head(d)
ggplot(d) + geom_point(aes(x = x, y = y, colour = group1)) + facet_grid(~group2)

```

For the exercises in class we'll use the `diamonds` dataset from the `ggplot2` package. This package and `cars` are commonly used data sets in `ggplot2` examples, you'll come across them a lot. 

`diamonds` is a rather large datset and ggplot can be kind of slow in these cases so we will trim it down to make our lives easier. Using the `sample` function we will randomly select 1000 rows. The default is to sample *without* replacement.    

```{r trimming-diamonds}
head(diamonds) # from the ggplot2 package
nrow(diamonds) # number of rows in the data frame...oof, that's a lot

set.seed(1) # 
diamonds2 <- diamonds[sample(nrow(diamonds), 1000), ]
nrow(diamonds2)
```

### Here's a basic plot with points to start with

```{r basic-plot}
ggplot(diamonds2, aes(x = carat, y = price)) + geom_point()
```

# My work flow with ggplot2
1. Make my data tidy (see `plyr` & `reshape2` lecture)
2. Open the [ggplot2 help page](http://docs.ggplot2.org/) 
3. Plot, plot, plot

> "Make 100 figures for every 1 you publish" -Marc Mangel

Let's start exploring our data by trying out at least 3 geoms with the diamonds2 dataset. geoms you might want to try: `geom_histogram`, `geom_bar`, `geom_violin`, `geom_point`, `geom_freqpoly` ...

### Histogram

```{r diamonds2-histogram1}
ggplot(diamonds2, aes(x = price)) + geom_histogram()

```

### Violin plot

```{r diamonds2-violin1}
ggplot(diamonds2, aes(x = cut, y = price)) + geom_violin()

```

### Frequency polygon

```{r diamonds2-freqpoly1}
ggplot(diamonds2, aes(x = price)) + geom_freqpoly()

```

Maybe you want to adjust the bind widths and x axis limits on your histogram

```{r diamonds2-histogram2}
ggplot(diamonds2, aes(x = price)) + geom_histogram(binwidth=100)

```

You can even add geoms on top of each other as layers. Try combining a histogram and frequency plot

```{r diamonds2-histogram-freqpoly}
ggplot(diamonds2, aes(x = price)) + geom_histogram(fill='red',color='black') + geom_freqpoly(size=2) # size adjusts the thickness of the line 
```

# Combine aesthetics to get the most out of your data

By combining the aesthetics of geom_point: x, y, alpha, colour,
shape, and size, see how many dimensions of the data you can display
simultaneously with aesthetics.

```{r diamonds-point-many-aes}
ggplot(diamonds2,aes(x=carat,y=price,color=color,size=price,alpha=depth)) + geom_point()
```


### Try expressing the same data with different aesthetics.

```{r diamonds2-point-shape}
ggplot(diamonds2,aes(x=carat,y=price,color=color,size=price,shape=cut)) + geom_point()
```

### What happens when you specify a continuous vs. a factor variable for an aesthetic like colour?

```{r diamonds2-point-color-continuous}
ggplot(diamonds2,aes(x=carat,y=price,color=price,size=price)) + geom_point()

```

# Use `facet_wrap` and `facet_grid` to group "like" data  
Try replacing one of the aesthetics with a facet_wrap.

```{r diamonds2-point-facet_wrap-cut}
ggplot(diamonds2,aes(x=carat,y=price,color=price,size=price,alpha=0.5)) + geom_point() + 
  facet_wrap(~cut)
```

Now use facet_grid with the diamonds dataset.

```{r diamonds2-point-facet_grid-clarity-cut}
ggplot(diamonds2,aes(x=carat,y=price,color=price,size=price,alpha=0.5)) + geom_point() + 
  facet_grid(clarity~cut)
```

## Scaling your data

By default all of the x and y axis limits are shared. If price is our dependent variable then the min and max limits on our y-axis will be determined by the min and max value of price in the **whole data set**. The `scales` option will let us scale our axes by the data in a facet'd group.

Try adjusting both scales independently and together 

### `free_x` 

```{r diamonds2-facet-scales-free_x}
ggplot(diamonds2,aes(x=carat,y=price,color=price,size=price,alpha=0.5)) + geom_point() + 
  facet_grid(clarity~cut,scale='free_x')

```

### `free_y`

```{r diamonds2-facet-scales-free_y}
ggplot(diamonds2,aes(x=carat,y=price,color=price,size=price,alpha=0.5)) + geom_point() + 
  facet_grid(clarity~cut,scale='free_y')
```
### `free`

```{r diamonds2-facet-scales-free}
ggplot(diamonds2,aes(x=carat,y=price,color=price,size=price,alpha=0.5)) + geom_point() + 
  facet_grid(clarity~cut,scale='free')
```

# Themes 

The default theme in ggplot2 isn't terrible. In fact it is a big improvement when compared to what you get in base plot, excel, JMP, etc. But if you are not happy with the default theme there are quite a few other 'pre-installed' themes


### default theme
```{r diamonds2-theme-default}
ggplot(diamonds2,aes(x=carat,y=price,color=clarity)) + geom_point() 

```

### grey theme
```{r diamonds2-theme-grey}
ggplot(diamonds2,aes(x=carat,y=price,color=clarity)) + geom_point() + theme_grey()

```

### black and white theme
```{r diamonds2-theme-bw}
ggplot(diamonds2,aes(x=carat,y=price,color=clarity)) + geom_point() + theme_bw()

```

## Even more themes
If you want to get crazy there are a whole bunch more themes in the package `ggthemes`

```{r}
install.packages('ggthemes')
library('ggthemes')
```

### Edward Tufte inspired theme
```{r diamonds2-theme-tufte}
ggplot(diamonds2,aes(x=carat,y=price,color=clarity)) + geom_point() + theme_tufte()

```

### Excel theme for the nostalgic
```{r diamonds2-theme-ImGoingToBeIll}
ggplot(diamonds2,aes(x=carat,y=price,color=clarity)) + geom_point() + theme_excel()

```

### Wall Street Journal
```{r diamonds2-theme-wsj}
ggplot(diamonds2,aes(x=carat,y=price,color=clarity)) + geom_point() + theme_wsj()

```


### Economist
```{r diamonds2-theme-economist}
ggplot(diamonds2,aes(x=carat,y=price,color=clarity)) + geom_point() + theme_economist()

```

# Customization within themes
within the theme function you can do all sorts of customization

Try making some modifications to our plot using `theme`. Try changing the color of the plot background and axis text. 

```{r diamonds2-custom-theme1}
ggplot(diamonds2,aes(x=carat,y=price,color=clarity)) + geom_point() + 
  theme(plot.background = element_rect(fil = "pink")) +
  theme(axis.text =element_text(color = 'blue'))

```

Now try modifying the color of the 'strip' box in a facet and moving the legend to the bottom.

```{r diamonds2-custom-theme2}
ggplot(diamonds2,aes(x=carat,y=price,color=clarity)) +   
  geom_point() + facet_wrap(~cut) +
  # white box with black outline for facet ("strip") labels
  theme(strip.background = element_rect(fill='white',color='black')) + 
  # move legend to the bottom (use 'none' to suppress the legend)
  theme(legend.position = 'bottom')
```

# Summary statistics and models

We usually want to do more than just look at our raw dat. The statistial tranformations allow us to transform, summarize, and fit simple models to our data. Try some of these options like `stat_bin`,`stat_density`,`stat_boxplot`, `stat_smooth`. 

### `stat_bin`
```{r diamonds2-stat_bin}
ggplot(diamonds2,aes(x=price)) + stat_bin()

```

### `stat_density`
```{r diamonds2-stat_density}
ggplot(diamonds2,aes(x=price)) + stat_density()
```

### `stat_boxplot`
```{r diamonds2-boxplot}
ggplot(diamonds2,aes(x=cut,y=price)) + stat_boxplot()
```

### `stat_smooth`
Try plotting price by carat and adding smoother. What happens when you add the points layer before the smooth layer? Try using a couple different types of smoothers (linear model, loess). What happens when you add and/or facets? 

```{r diamonds2-stat_smooth}
ggplot(diamonds2,aes(x=carat,y=price)) + stat_smooth(method='loess') + geom_point()
```

###  `stat_function`
```{r stat_function-normal-curve}
ggplot(data.frame(x = c(-5, 5)), aes(x)) + stat_function(fun = dnorm)
```
 
# Plotting Results
`ggplot` can also be useful for plotting results
Use ddply to get the mean and standard deviation price for each diamond cut quality

```{r ddply-diamonds2}
library(plyr)
diamonds.price <- ddply(diamonds2,'cut',summarise,
             mean = mean(price),
             sd = sd(price))

```

### Plotting the results

```{r diamonds-pointrange}
ggplot(diamonds.price,aes(x=cut,y=mean,ymin=mean-sd,ymax=mean+sd)) + 
  geom_pointrange() + geom_point(color='red') 
```

### Alternatively using errorbars
```{r diamonds-errorbar}
ggplot(diamonds.price,aes(x=cut,y=mean,ymin=mean-sd,ymax=mean+sd)) + 
  geom_errorbar(width=0.5) + geom_point(color='red') 
```

### Flipping our coordinates
```{r diamonds-pointrange-coordflip}
ggplot(diamonds.price,aes(x=cut,y=mean,ymin=mean-sd,ymax=mean+sd)) + 
  geom_pointrange() + geom_point(color='red') + coord_flip() 
```

### Add a reference line
```{r diamonds-pointrange-coordflip-refline}
ggplot(diamonds.price,aes(x=cut,y=mean,ymin=mean-sd,ymax=mean+sd)) + 
  geom_hline(y=3000,linetype=2,size=2) + geom_pointrange() + geom_point(color='red') + coord_flip() + ylab('price')

# remember that we have 'flipped' the coordinates so our hline will actually make a verticl line and will require a y intercept 
```

